---
title: "Homework 4"
author: Faustine, Eden, Shaoji, Yumemichi
output: html_document
---

> "La Quinta is Spanish for next to Denny’s" - Mitch Hedberg

Our group scraped location data from La Quinta and Denny's websites to see if this joke has a kernal of truth in it. 


### Task 1 - Scraping La Quinta

`get_lq.R` creates and save the html files of each La Quinta hotel. First we obtained the url of hotel listing page from the La Quinta website. This url page contains all the web links for each individual hotel. 

Then we extracted individual hotel page from the url. Installing R package `rvest` (for processing and manipulation of html data), we used `html_read()` to read html data from the url of individual html pages. Then `html_nodes()` specifies the nodes as hotel listings from the html document using CSS selectors, and `html_attr('href')` extracts all of each hotel's attribute by its destination, removing the missing values. 

For the next step, we downloaded the resulting html files of each La Quinta location to the created folder `"data/lq"`. Looping over url of each hotel obtained by the previous step, `download.file()` downloads the html file from the Internet to the directory with the location name concatenated. We set `quiet = TRUE` to suppress status messages and the progress bar.

We moved on to parse the La Quinta's html files by `parse_lq.R`. First we loaded the sequence of html files in the created directory using `files()` and initialzed the results that hold all the information as a list. Then we extracted the desirable information using a for loop. Inside the for loop, we read every html file and used `html_nodes()` to extract the basic information about the hotel, including address, Phone and Fax information. 

Using SelectorGadget and piping, we used `html_text()` to extract the text of the information as a vector of characters, and then used `str_split()` to split the information to Address, Phone number and Fax number in different lines; we split the vectorby "\n", the separator in the vector. After the split, we unlisted the object and use `str_trim()` to remove the extra space. Same method of parsing was applied to obtain information about the location name, room numbers, floor numbers and amenities/services. 

For latitude and longitude, we used regular expressions to locate the numbers in the html file  of the displayed Google minimap of the hotel information website. For the availabilities of Internet acees and swimming pool, we first obtained the entire information of all amenities and sevices and then used `str_detect() ` to find if there is swimming pool and internet acces in the given hotel; `any()` returns the result as binary variables. 

We put all the information obtained from the loop in a data frame; each row represents the entire information (location name, address, phone, fax, number of rooms, latitude, longtitude, number of floors, swimming pool and internet access) of each hotel. 

To make future tasks easier, we added an extra column `state` to the dataframe `hotels` to indicate which state homes each of the La Quinta hotel. First we initialize an empty sequence to hold the information of all the hotels (902 in total). Then we filled in the matrix with a loop: in each interation, we extracted and unlisted the address of each hotel, with commas removed, and then capture the state abbrev nested in the zip code.

Finally, we saved the new dataframe `hotels` to the created directory `\data` with the data file named as `lq.Rdata`. 


### Task 2 - Scraping Denny's

The location of each Denny's was obtained by interacting with a third-party location service - this is the service that powers the location map and search feature on the Denny's website. In this case we queried the location provider "Where 2 Get It".

The `get_dennys` script takes the form of an xml request to the where2getit domain. In that request, we first translated the url to the form that was more interpreble. We found three important xml calls: `zipcode`, `radius`, and `limit`. 

The `zipcode` value is a US zipcode that gives us a center point. The `radius` is the maximum distance in miles that the locator looks out from the zipcode. Limit gives the maxiumum number of resturants to return. 

We set the `zipcodes` strategically to cover the entire US. We picked locations in the East and West coast to cover the continetal US. We also picked locations in Alaska and Hawaii to cover all 50 states. The four cities we choose were Washington DC, Los Angeles California, Honolulu HI, and Fairbanks AK. Zipcodes for these cities were found on Wikipedia. 

We wanted the `radius` value to be the smallest values that cover all the locations. We increased the radius until the number of locations stopped increasing. For the four cities we choose, that number happened to be `radius = 5000`. Finally, we set the limit to be `10000`, which is much higher than the number of Denny's locations so that we could be sure we aren't missing some values. 

Finally, we downloaded the resulting xml files with the infomation of Denny's locations in the `data/dennys` folder.

We parsed the Denny's xml files similarly to the La Quinta data. We read the xml file using `rvest` and used the `xml_text()` to extract the text. The location information is in the `<latitude>` and `longitude` node. We also parsed the state data. By looking at the state data we see that we got all the Denny's from 51 states (50 states plus Washington DC).



### Task 3 - Distance Analysis

To estimate the veracity of the statement on the general trend of distances between La Quinta and Denny's, we used *mapping of locations* (ggplot) and *histogram* of number of nearby Denny's for a fixed La Quinta. We also calculated the *proportions* of La Quinta's that have at least one Denny's within different radius.


## Mapping of locations 
First we need to rebuild dataframes removing characteristics other than lattitude, longtitude and state. Furthermore, we are only interested in the case of U.S., so we removed the rows (La Quinta or Denny's) without nonempty characteristis `state`. The dataframes were updated as `lqLoc` and `dennysLoc`. 
```{r}
load('data/lq.Rdata')
load('data/dennys.Rdata')
library('dplyr')

# load packages for visualization:
library('maps')
library('datasets')
library('ggplot2')


# Build data frames for La Quinta and Denny's that include state, lattitude and longitude
lqLoc = data.frame(State = hotels$state, 
                   Lattitude = as.numeric(hotels$lat), 
                   Longitude = as.numeric(hotels$long))
dennysLoc = data.frame(State = dennys$state, 
                   Lattitude = as.numeric(dennys$lat), 
                   Longitude = as.numeric(dennys$long))

#  Remove the denny's and La Quinta with no State, which means they are outside of US
lqLoc = lqLoc[!is.na(lqLoc$State),]
dennysLoc = dennysLoc[dennysLoc$State != "",]
```

Note that it makes sense to compare the proximity of Denny's and La Quinta in a state where both of them exists. Therefore we obtain `dennys_commonStates` and `lq_commonStates` by filtering the common states (by `unique()` and `intersect()`), and then subsetting the dataframe whose charactersitics of `State` falls in the sequence of common states. Beforehand we need to reordering the dataframe, Then we won't worry about comparing a Denny's in Minnesota with a La Quinta in North Carolina that leads to large bias in the final result. 
```{r}
# Re-order the data frame by state
lqLoc = lqLoc[order(lqLoc$State),]
dennysLoc = dennysLoc[order(dennysLoc$State),]

# Find the states where both Denny's and La Quinta exists
lq_states = unique(lqLoc$State)
dennys_states = unique(dennysLoc$State)
common_states = intersect(lq_states, dennys_states)

# Selecting Denny's and La Quinta that exist in states having both of them
dennys_commonStates = dennysLoc[dennysLoc$State %in% common_states,]
lq_commonStates = lqLoc[lqLoc$State %in% common_states,]
```

Then we plot all the locations of Denny's and La Quinta on the US map (including Alaska). First we make a new data frame that stores all La Quinta and Denny's information for ggplot, and then create a data frame that stores information of only US states. Then we use `ggplot()` to show each location of the two types of business.
```{r}
# Make a new data frame that stores all La Quinta and Denny's information for ggplot
lq_melt = data.frame(Type = rep("La Quinta", nrow(lq_commonStates))) %>%
  cbind(lq_commonStates)
dennys_melt = data.frame(Type = rep("Denny's", nrow(dennys_commonStates))) %>%
  cbind(dennys_commonStates)
lq_dennys = rbind(lq_melt,dennys_melt)
lq_dennys$Type = as.factor(lq_dennys$Type)

# Create a data frame that stores information of all US states
world = map_data("world")
Alaska = world[world$subregion == "Alaska",] %>%
  .[!is.na(.$subregion),]
Alaska$region = Alaska$subregion
Alaska$subregion = rep(NA, nrow(Alaska))
all_states = rbind(map_data("state"), Alaska)

# Do the plot:
ggplot() + 
  geom_polygon(data = all_states, aes(x=long, y=lat, group = group), colour="white", fill="blue") +
  xlim(-170, -65) + 
  geom_point(data = lq_dennys, aes(x=Longitude, y=Lattitude, group=Type, color=Type), cex = 0.6)
```
From the graph we can see that Denny's and La Quintas have similar location distribution nationwide.


##Histogram of Number of Nearby Denny's for a La Quinta

To take a further look at the proximity of these two business, a histogram of total number of nearby Denny's would be useful.

First we need to initialize an empty matrix with dimension as the number of Denny's (row) by the number of La Quintas (column). Then we fill the matrix by fixing one La Quinta (subsetting each of the sequence by the index) and compute the distance of each Denny's with this particular la Quinta; each iteration generates a result by column. The distance formula is the spherical law of cosines. 

Going through the loop we obtained a full matrix of distances. For example, the *i,j* entry of the matrix *distMat* denotes the distance between *ith Denny's* in the dataframe and *jth La Quinta*. 

Then we count the number of Denny's within 5 miles in distance for each La Quinta.
```{r}
# formula: spherical law of cosines
# reference: http://www.movable-type.co.uk/scripts/latlong.html

# Build a distance matrix for denny's and La Quinta, with the rows representing Denny's and columns representing La Quinta
distMat = matrix(rep(0,nrow(dennys_commonStates)*nrow(lq_commonStates)), 
                 nrow = nrow(dennys_commonStates),
                 ncol = nrow(lq_commonStates))
for(i in 1:nrow(lq_commonStates)) {
  lat1 = rep(lq_commonStates$Lattitude[i], nrow(dennys_commonStates))
  lon1 = rep(lq_commonStates$Longitude[i], nrow(dennys_commonStates))
  lat2 = dennys_commonStates$Lattitude
  lon2 = dennys_commonStates$Longitude
  # fill the matrix by column (fixing the La Quinta)
  distMat[,i] = acos(sin(lat1*pi/180)*sin(lat2*pi/180)+
           cos(lat1*pi/180)*cos(lat2*pi/180)*cos(lon2*pi/180-lon1*pi/180)) * 3959
}

counts = c() 
for(i in 1:nrow(lq_commonStates)) {
  counts = c(counts, sum(distMat[,i] < 5))   #concatenate the sequence of counts 
}
```


Finally, we plot the histogram of the number of nearby Denny's for each La Quinta:
```{r}
hist(counts, main="number of nearby Denny's for a La Quinta", xlab = "numbers", ylab = "")
```



## "Proportion analysis" --- proportion of La Quinta's that neighbors a Denny's within different choices of radius.

The general method of this analysis is as follows:

1) Initialize an empty sequence with length of La Quintas. 

2) Iterating over the for loop, compute the total number of La Quinta that neighbors a Denny's within the specified distance

3) Compute the ratio of this number of La Quinta to the total number of La Quintas. 

According to the following results, the La Quintas and Denny's indeed seem to be loyal neighbors. 

**1) Proportion of La Quintas that have at least one Denny's within 1 mile is around 26%**
```{r}
within_1mile = rep(NA, nrow(lq_commonStates))
for(j in 1:nrow(lq_commonStates)) {
  within_1mile[j] = sum(distMat[,j] < 1)
}
sum(within_1mile > 0) / length(within_1mile)
```
  
**2) Proportion of La Quintas that have at least one Denny's within 5 miles is around 60%**
```{r}
within_5miles = rep(NA, nrow(lq_commonStates))
for(j in 1:nrow(lq_commonStates)) {
  within_5miles[j] = sum(distMat[,j] < 5)
}
sum(within_5miles > 0) / length(within_5miles)
```
  
**3) Proportion of La Quintas that have at least one Denny's within 10 miles is around 72%**
```{r}
within_10miles = rep(NA, nrow(lq_commonStates))
for(j in 1:nrow(lq_commonStates)) {
  within_10miles[j] = sum(distMat[,j] < 10)
}
sum(within_10miles > 0) / length(within_10miles)
```
  
**4) Proportion of La Quintas that have at least one Denny's within 20 miles is around 80%**
```{r}
within_20miles = rep(NA, nrow(lq_commonStates))
for(j in 1:nrow(lq_commonStates)) {
  within_20miles[j] = sum(distMat[,j] < 20)
}
sum(within_20miles > 0) / length(within_20miles)
```
  

## Further Discussion:

To further verify the joke about these two neighbors, more analysis is necessary. We can introduce some blocks to the location such as metro area, along the highway or countryside. More comparisons with Denny's should be done as well. For example, we could go through the same procedure and compare the distance between West Bestern, Motel 6 or Quality Inn (or other peer budget hotels) and Denny's. Conversely, we should also compare the proximity of Perkins, Dominos or Piza Hut to La Quinta's and see if any of the casual restaurant favors La Quinta in particular. 





